---
title: "Lab3A"
author: "Abigail Booth"
date: "9/7/2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r,eval=FALSE}
 download.file(url="https://ndownloader.figshare.com/files/2292169",
              destfile = "data/portal_data_joined.csv")
```

```{r}
surveys <- read.csv("data/portal_data_joined.csv")
head(surveys)
```

Looking at only the first few lines of data
```{r}
head(surveys)
```
```{r}
View(surveys)
```
To see the structure of the data frame
```{r}
str(surveys)
```

### Challenge: Inspecting data.frame objects
 - The class of the object surveys is data.frame, as is stated in the top left corner of the output from str(surveys).
 - There are 34,786 rows and 13 columns in the object. This information can be found through the "34786 obs. of 13 variables" from str(surveys), or using nrow and ncol commands. 
```{r}
nrow(surveys)
ncol(surveys)
```
 - There are 48 species that were recorded during these surveys. This comes from the str(surveys) command where it states that within the column species_id, there are 48 levels.
 
### Indexing and subsetting data frames
```{r}
# first element in the first column of the data frame (as a vector)
surveys[1, 1]   
```
```{r}
# first element in the 6th column (as a vector)
surveys[1, 6]
```
```{r}
# first column of the data frame (as a vector)
surveys[, 1]
```
```{r}
# first column of the data frame (as a data.frame)
surveys[1]
```
```{r}
# first three elements in the 7th column (as a vector)
surveys[1:3, 7] 
```
```{r}
# the 3rd row of the data frame (as a data.frame)
surveys[3, ] 
```
```{r}
# equivalent to head_surveys <- head(surveys)
head_surveys <- surveys[1:6, ]
```
```{r}
surveys[-(7:34786), ] # Equivalent to head(surveys)
```
```{r}
surveys["species_id"]       # Result is a data.frame
```
```{r}
surveys[, "species_id"]     # Result is a vector
```
```{r}
surveys[["species_id"]]     # Result is a vector
```
```{r}
surveys$species_id          # Result is a vector
```
### Challenge: Indexing and subsetting data frames
1. 
```{r}
surveys_200 <- surveys[200, ]
surveys_200
```
2. 
```{r}
num_rows <- nrow(surveys)
num_rows
```
```{r}
surveys[34786, ]
```
```{r}
tail(surveys)
```
```{r}
surveys_last <- surveys[num_rows, ]
surveys_last
```
3. 
```{r}
surveys_middle <- surveys[nrow(surveys)/2, ]
surveys_middle
```
4. 
```{r}
first_six_rows <- surveys[-(7:nrow(surveys)), ]
first_six_rows
```
### Factors
```{r}
sex <- factor(c("male", "female", "female", "male"))
levels(sex)
nlevels(sex)
```
```{r}
sex # current order
sex <- factor(sex, levels = c("male", "female"))
sex # after re-ordering
```
```{r}
as.character(sex)
```
```{r}
year_fct <- factor(c(1990, 1983, 1977, 1998, 1990))
as.numeric(year_fct)               # Wrong! And there is no warning...
as.numeric(as.character(year_fct)) # Works...
as.numeric(levels(year_fct))[year_fct]    # The recommended way.
```
```{r}
## bar plot of the number of females and males captured during the experiment:
plot(as.factor(surveys$sex))
```
```{r}
sex <- factor(surveys$sex)
head(sex)
levels(sex)
levels(sex)[1] <- "undetermined"
levels(sex)
head(sex)
```
### Challenge: Renaming factors
```{r}
levels(sex)[2:3] <- c("female", "male")
levels(sex)
```
```{r}
sex <- factor(sex, levels = c("female", "male", "undetermined"))
plot(sex)
```

### Using stringsAsFactors = FALSE
```{r}
## Compare the difference between our data read as `factor` vs `character`.
surveys <- read.csv("data/portal_data_joined.csv", stringsAsFactors = TRUE)
str(surveys)
```
```{r}
surveys <- read.csv("data/portal_data_joined.csv", stringsAsFactors = FALSE)
str(surveys)
```
```{r}
## Convert the column "plot_type" into a factor
surveys$plot_type <- factor(surveys$plot_type)
```
### Challenge: Using stringsAsFactors = FALSE
1. In animal, there were no quotes around the types of animals. In feel, there were not enough entries to be consistent in the number of columns. In weight, there was no comma between 8 and 1.1.

Corrected: 
```{r}
animal_data <- data.frame(
          animal = c("dog", "cat", "sea cucumber", "sea,urchin"),
          feel = c("furry", "squishy", "spiny", "scaly"),
          weight = c(45, 8, 1.1, 0.8))
```
2. I would predict that country, climate, northern_hemisphere, and has_kangaroo are factors because they are all categorical and temperature is numeric. 
```{r}
country_climate <- data.frame(
       country = c("Canada", "Panama", "South Africa", "Australia"),
       climate = c("cold", "hot", "temperate", "hot/temperate"),
       temperature = c(10, 30, 18, "15"),
       northern_hemisphere = c(TRUE, TRUE, FALSE, "FALSE"),
       has_kangaroo = c(FALSE, FALSE, FALSE, 1))
str(country_climate)
```
The classes are not what I would have expected, but since has_kangaroo is the only column to have no quotations and a number, it makes sense that it is a numeric. If stringsAsFactors = FALSE would have converted all of the factors to character vectors because there is text. To make sure eack column had the accurate data type, the entries would need to be checked to confirm that the intended classes were correct. More specifically, the quotes would need to be removed from 15 in climate, the FALSE would need to be taken out of quotes in northern_hemisphere, and the 1 in has_kangaroo would most likely need to be changed to TRUE. 

### Formatting Dates
```{r}
str(surveys)
```
```{r}
library("lubridate")
```
```{r}
my_date <- ymd("2015-01-01")
str(my_date)
```
```{r}
# sep indicates the character to use to separate each component
my_date <- ymd(paste("2015", "1", "1", sep = "-")) 
str(my_date)
```
```{r}
paste(surveys$year, surveys$month, surveys$day, sep = "-")
```
```{r}
ymd(paste(surveys$year, surveys$month, surveys$day, sep = "-"))
```
```{r}
surveys$date <- ymd(paste(surveys$year, surveys$month, surveys$day, sep = "-"))
str(surveys) # notice the new column, with 'date' as the class
```
```{r}
summary(surveys$date)
```
```{r}
missing_dates <- surveys[is.na(surveys$date), c("year", "month", "day")]
head(missing_dates)
```

