---
title: "lab4_alb.Rmd"
author: "Abigail Booth"
date: "September 14, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```
```{r}
report_03_11_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-11-2020.csv")) %>%
  rename(Country_Region = "Country/Region", Province_State = "Province/State")
```

```{r}
str(report_03_11_2020)
```
```{r}
View(report_03_11_2020)
```
```{r}
library(DT)
```
```{r}
datatable(report_03_11_2020)
```
### Exercises (part 1)
1. The report for 9/13/20 is significantly longer with 3,954 rows and 14 columns in comparison to the 230 rows and 8 columns on 3/13/20. Therefore, many more countries and provinces are included in the later data. A few of the column titles are changed, such as "Latitude" was shortened to "Lat" in the 9/13 data. Also, the report for 9/13/20 has expanded to include information on active cases, combined key, incidence rate, and case fatality ratios with FIPS and Admin2.
```{r}
report_03_13_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-13-2020.csv")) %>% 
  rename(Country_Region = "Country/Region", Province_State = "Province/State")
```
```{r}
report_09_13_2020 <-  read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-13-2020.csv"))
```
```{r}
str(report_03_13_2020)
```

```{r}
str(report_09_13_2020)
```
```{r}
View(report_03_13_2020)
View(report_09_13_2020)
```
2. 
```{r}
count_09_13_2020 <- report_09_13_2020 %>%
  group_by(Country_Region) %>%
  summarize(Confirmed = sum(Confirmed),
            Deaths = sum(Deaths),
            Recovered = sum(Recovered),
            Active = sum(Active))
count_09_13_2020
```
3. 
```{r}
US_confirmed <- report_09_13_2020 %>%
  filter(!is.na(Province_State)) %>% 
  filter(Country_Region == "US") %>%
  group_by(Province_State) %>% 
  summarize(confirmed_cases = sum(Confirmed))
US_confirmed
```
4. 
```{r}
top_10_deaths <- report_09_13_2020 %>%
  group_by(Country_Region) %>%
  summarize(max_deaths = max(Deaths)) %>%
  arrange(desc(max_deaths)) %>%
  slice(1:10)
top_10_deaths
```
### Exercises (Part 2)
5. 
```{r}
report_03_12_2020 <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-12-2020.csv") %>% 
  rename(Country_Region = "Country/Region", Province_State = "Province/State")
```
```{r}
highest_cases_and_deaths_03_12 <- report_03_12_2020 %>%
  group_by(Country_Region) %>%
  summarize(confirmed_cases = max(Confirmed),
            deaths = max(Deaths)) %>%
  arrange(desc(confirmed_cases)) %>%
  slice(1:10)
highest_cases_and_deaths_03_12
```

Countries with Highest Deaths - March 2020

```{r}
ggplot(data = highest_cases_and_deaths_03_12, mapping = aes(x = deaths, y = Country_Region)) +
  geom_point()
```

Countries with Highest Confirmed Cases - March 2020

```{r}
ggplot(data = highest_cases_and_deaths_03_12, mapping = aes(x = confirmed_cases, y = Country_Region)) +
  geom_point()
```

```{r}
report_09_12_2020 <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-12-2020.csv")
```
```{r}
highest_cases_and_deaths_09_12 <- report_09_12_2020 %>%
  group_by(Country_Region) %>%
  summarize(confirmed_cases = max(Confirmed),
            deaths = max(Deaths)) %>%
  arrange(desc(confirmed_cases)) %>%
  slice(1:10)
highest_cases_and_deaths_09_12
```

Countries with Highest Deaths - September 2020

```{r}
ggplot(data = highest_cases_and_deaths_09_12, mapping = aes(x = deaths, y = Country_Region)) +
    geom_point()
```

Countries with Highest Confirmed Cases - September 2020

```{r}
ggplot(data = highest_cases_and_deaths_09_12, mapping = aes(x = confirmed_cases, y = Country_Region)) +
    geom_point()
```

6. 
```{r}
US_highest_cases_and_deaths_03_12 <- report_03_12_2020 %>%
  filter(Country_Region == "US") %>%
  group_by(Province_State) %>%
  summarize(confirmed_cases = max(Confirmed),
            deaths = max(Deaths)) %>%
  arrange(desc(confirmed_cases)) %>%
  slice(1:10)
US_highest_cases_and_deaths_03_12
```

States with Highest Deaths - March 2020

```{r}
ggplot(data = US_highest_cases_and_deaths_03_12, mapping = aes(x = deaths, y = Province_State)) +
  geom_bar(stat = "identity")
```

States with Highest Confirmed Cases - March 2020

```{r}
ggplot(data = US_highest_cases_and_deaths_03_12, mapping = aes(x = confirmed_cases, y = Province_State)) +
  geom_bar(stat = "identity")
```
```{r}
US_highest_cases_and_deaths_09_12 <- report_09_12_2020 %>%
  filter(Country_Region == "US") %>%
  group_by(Province_State) %>%
  summarize(confirmed_cases = max(Confirmed),
            deaths = max(Deaths)) %>%
  arrange(desc(confirmed_cases)) %>%
  slice(1:10)
US_highest_cases_and_deaths_09_12
```

States with Highest Deaths - September 2020

```{r}
ggplot(data = US_highest_cases_and_deaths_09_12, mapping = aes(x = deaths, y = Province_State)) +
  geom_bar(stat = "identity")
```

States with Highest Confirmed Cases - September 2020

```{r}
ggplot(data = US_highest_cases_and_deaths_09_12, mapping = aes(x = confirmed_cases, y = Province_State)) +
  geom_bar(stat = "identity")
```