---
title: "Lab_9"
author: "Abigail Booth"
date: "October 26, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Matching patterns w/ regular expressions

```{r, message=FALSE, warning=FALSE}
library(stringr)
library(tidyverse)
```

# Basic matches

```{r}
x <- c("apple", "banana", "pear")
str_view(x, "an")
```
```{r}
str_view(x, ".a.")
```
```{r}
# To create the regular expression, we need \\
dot <- "\\."

# But the expression itself only contains one:
writeLines(dot)
```
```{r}
# And this tells R to look for an explicit .
x <- c("abc", "a.c", "bef")
str_view(x, "a\\.c")
```
```{r}
x <- "a\\b"
writeLines(x)
str_view(x, "\\\\")
```

# Exercises

1. The strings don't match because \ is the literal regular expression and "\" is an escape character meaning that it just goes down to the next character. "\\" codes for the escape character and "\\\\" matches a literal \ and then escapes to the following character. 

2. You would match the sequence "'\ with:
```{r}
x <- "\"'\\"
writeLines(x)
str_view(x, "\"'\\\\")
```

# Anchors

```{r}
x <- c("apple", "banana", "pear")
str_view(x, "^a")
```
```{r}
str_view(x, "a$")
```
```{r}
x <- c("apple pie", "apple", "apple cake")
str_view(x, "apple")
```
```{r}
str_view(x, "^apple$")
```

# Exercises

1. You would match the literal string "$^$" with:
```{r}
str_view("$^$", "^\\$\\^\\$$")
```
2.
```{r}
# find words that start w/ y
str_view(stringr::words, "^y", match=TRUE)
```
```{r}
# find words that end with x
str_view(stringr::words, "x$", match=TRUE)
```
```{r}
# find words that are exactly 3 letters long
str_view(stringr::words, "^...$", match=TRUE)
```
```{r}
# find words that have 7 letters or more
str_view(stringr::words, ".......", match=TRUE)
```

# Character classes and alternatives

```{r}
# Look for a literal character that normally has special meaning in a regex
x <- c("abc", "a.c", "a*c", "a c")
str_view(x, "a[.]c")
```
```{r}
str_view(x, ".[*]c")
```
```{r}
str_view(x, "a[ ]")
```
```{r}
x <- c("grey", "gray")
str_view(x, "gr(e|a)y")
```
# Exercises

1. 
```{r}
# find words that start w/ a vowel
str_view(stringr::words, "^(a|e|i|o|u)", match=TRUE)
```
```{r}
# find words that only contain consonants
str_view(stringr::words, "(a|e|i|o|u)", match=FALSE)
```
```{r}
# find words that end w/ ed, but not eed
str_view(stringr::words, "[^e]ed$", match=TRUE)
```
```{r}
# find words that end w/ ing or ise
str_view(stringr::words, "(ing|ise)$", match=TRUE) 
```
2.
```{r}
# i before e except after c
length(str_subset(stringr::words, "(cei|[^c]ie)"))
# e before i except after c
length(str_subset(stringr::words, "(cie|[^c]ei)"))
```
3.
```{r}
length(str_subset(stringr::words, "q[^u]"))
```
In the case of all the words in stringr::words, all of themhave q followed by u.

# Repetition

```{r}
x <- "1888 is the longest year in Roman numerals: MDCCCLXXXVIII"
str_view(x, "CC?")
```
```{r}
str_view(x, "CC+")
```
```{r}
str_view(x, 'C[LX]+')
```

```{r}
str_view(x, "C{2}")
```
```{r}
str_view(x, "C{2,}")
```
```{r}
str_view(x, "C{2,}")
```
```{r}
str_view(x, 'C{2,3}?')
```
```{r}
str_view(x, 'C[LX]+?')
```

# Exercises

1. The equivalent of ? is {0, 1}, + is {1,}, and * is {0,}

2. 
  * ^.*$ is any string
  * "\\{.+\\}" is any string with {} surrrounding at least one character
  * \d{4}-\d{2}-\d{2} is any 4 numbers followed by a hyphen followed by any two numbers followed by a hyphen followed by any 2 numbers
  * "\\\\{4}" is 4 backslashes

3.
```{r}
# find words that start w/ 3 consonants
str_view(stringr::words, "^[^a|e|i|o|u]{3}", match=TRUE)
```
```{r}
# find words that have 3 or more vowels in a row
str_view(stringr::words, "[a|e|i|o|u]{3,}", match=TRUE)
```
```{r}
# find words that have 2 or more vowel-consonant pairs in a row
str_view(stringr::words, "([a|e|i|o|u][^a|e|i|o|u]){2,}", match=TRUE)
```

# Grouping and backreferences

```{r}
str_view(fruit, "(..)\\1", match = TRUE)
```

# Exercises

1. 
  * (.)\1\1 matches the same character 3 consecutive times
  * "(.)(.)\\2\\1" matches a group of 2 characters followed by those characters in the opposite order
  * (..)\1 matches any two characters that are repeated at least once
  * "(.).\\1.\\1" matches a character followed by any character, the first character followed by any character followed by the first character.
  * "(.)(.)(.).*\\3\\2\\1" matches 3 characters followed by 0 or more characters followed by the first 3 characters in a reversed order.
  
2. 
```{r}
# find words that start and end w/ same character
str_view(stringr::words, "^(.)((.*\\1$)|(\\1?$))", match=TRUE)
```
```{r}
# find words that contain a repeated pair of letters
str_view(stringr::words, "(..).*\\1", match=TRUE)
```
```{r}
# find words that contain one letter repeated in at least 3 places
str_view(stringr::words, "([a-z]).*\\1.*\\1", match=TRUE)
```

# Detect matches

```{r}
x <- c("apple", "banana", "pear")
str_detect(x, "e")
```
```{r}
# How many common words start with t?
sum(str_detect(words, "^t"))
```

```{r}
# What proportion of common words end with a vowel?
mean(str_detect(words, "[aeiou]$"))
```
```{r}
# Find all words containing at least one vowel, and negate
no_vowels_1 <- !str_detect(words, "[aeiou]")
```
```{r}
# Find all words consisting only of consonants (non-vowels)
no_vowels_2 <- str_detect(words, "^[^aeiou]+$")
identical(no_vowels_1, no_vowels_2)
```
```{r}
words[str_detect(words, "x$")]
```

```{r}
str_subset(words, "x$")
```
```{r}
df <- tibble(
  word = words, 
  i = seq_along(word)
)
df %>% 
  filter(str_detect(word, "x$"))
```
```{r}
x <- c("apple", "banana", "pear")
str_count(x, "a")
```
```{r}
# On average, how many vowels per word?
mean(str_count(words, "[aeiou]"))
```
```{r}
df %>% 
  mutate(
    vowels = str_count(word, "[aeiou]"),
    consonants = str_count(word, "[^aeiou]")
  )
```

Note that matches never overlap. For example, in `"abababa"`, how many times will the pattern `"aba"` match? Regular expressions say two, not three:

```{r}
str_count("abababa", "aba")
```
```{r}
str_view_all("abababa", "aba")
```

# Exercises

1. 
```{r}
# one regular expression
str_view(stringr::words, "^x|x$", match=TRUE)
```
```{r}
# multiple str_detect()
words[str_detect(words, "^x")]
words[str_detect(words, "x$")]
```
```{r}
# one regular expression
str_view(stringr::words, "^[a|e|i|o|u].*[^(a|e|i|o|u)]$", match=TRUE) 
```
```{r}
# multiple str_detect()
words[str_detect(words, "^[a|e|i|o|u]")] %>% head()
words[str_detect(words, "[^(a|e|i|o|u)]$")] %>% head()
```

```{r}
words[str_detect(words, "a") &
  str_detect(words, "e") &
  str_detect(words, "i") &
  str_detect(words, "o") &
  str_detect(words, "u")]
```
There're no words that contain at least one of each different vowel.

# Extract matches

```{r}
length(sentences)
```
```{r}
head(sentences)
```
```{r}
colours <- c("red", "orange", "yellow", "green", "blue", "purple")
colour_match <- str_c(colours, collapse = "|")
colour_match
```
```{r}
has_colour <- str_subset(sentences, colour_match)
matches <- str_extract(has_colour, colour_match)
head(matches)
```
```{r}
# Test regex
has_colour
```
```{r}
# select sentences w/ more than 1 match
more <- sentences[str_count(sentences, colour_match) > 1]
str_view_all(more, colour_match)
```
```{r}
str_extract(more, colour_match)
```
```{r}
str_extract_all(more, colour_match)
```
```{r}
str_extract_all(more, colour_match, simplify = TRUE)
```
```{r}
x <- c("a", "a b", "a b c")
str_extract_all(x, "[a-z]", simplify = TRUE)
```

# Exercises

1. 
```{r}
colors_fixed <- c("red", "orange", "yellow", "green", "blue", "purple")
color_match_fixed <- str_c("\\b(", str_c(colors_fixed, collapse = "|"), ")\\b")
more <- sentences[str_count(sentences, color_match_fixed) > 1]
str_view_all(more, color_match_fixed)
```
2.
```{r}
# extract first word from each sentence
str_extract(sentences, "[A-Za-z][A-Za-z']*") %>%  head()
```
```{r}
# Extract all words ending in ing
ing <- "\\b[a-z]+ing\\b"
sent_w_ing <- str_detect(sentences, ing)
unique(unlist(str_extract_all(sentences[sent_w_ing], ing))) %>%  head()
```

# Grouped matches

```{r}
noun <- "(a|the) ([^ ]+)"
has_noun <- sentences %>%
  str_subset(noun) %>%
  head(10)
has_noun %>% 
  str_extract(noun)
```
```{r}
has_noun %>% 
  str_match(noun)
```
```{r}
tibble(sentence = sentences) %>% 
  tidyr::extract(
    sentence, c("article", "noun"), "(a|the) ([^ ]+)", 
    remove = FALSE
  )
```

# Exercises

1. 
```{r}
after_num <- "\\b(one|two|three|four|five|six|seven|eight|nine|ten) ([^ ]+)"

sentences[str_detect(sentences, after_num)] %>%
  str_extract(after_num)
```
2. 
```{r}
contract <- "([A-Za-z]+)'([A-Za-z]+)"

sentences[str_detect(sentences, contract)] %>%
  str_extract(contract) %>%
  str_split("'")
```

# Replacing matches

```{r}
x <- c("apple", "pear", "banana")
str_replace(x, "[aeiou]", "-")
```
```{r}
str_replace_all(x, "[aeiou]", "-")
```
```{r}
x <- c("1 house", "2 cars", "3 people")
str_replace_all(x, c("1" = "one", "2" = "two", "3" = "three"))
```
```{r}
sentences %>% 
  str_replace("([^ ]+) ([^ ]+) ([^ ]+)", "\\1 \\3 \\2") %>% 
  head(5)
```

# Exercises

1.
```{r}
slashes <- c("yes/no/maybe")
str_replace_all(slashes, "/", "\\\\")
```

# Splitting

```{r}
sentences %>%
  head(5) %>% 
  str_split(" ")
```
```{r}
"a|b|c|d" %>% 
  str_split("\\|") %>% 
  .[[1]]
```
```{r}
sentences %>%
  head(5) %>% 
  str_split(" ", simplify = TRUE)
```
```{r}
fields <- c("Name: Hadley", "Country: NZ", "Age: 35")
fields %>% str_split(": ", n = 2, simplify = TRUE)
```
```{r}
x <- "This is a sentence.  This is another sentence."
str_view_all(x, boundary("word"))
```
```{r}
str_split(x, " ")[[1]]
```
```{r}
str_split(x, boundary("word"))[[1]]
```

# Exercises

1. 
```{r}
str_split("apples, pears, and bananas", boundary("word"))[[1]]
# get rid of and
str_split("apples, pears, and bananas", ", +(and +)?")[[1]]
```
2. It's better to split up by boundaries rather than spaces because there could be commas and other forms of punctuation (other than apostrophies), which the " " couldn't account for. 

3. Splitting with an empty string splits the string into individual characters.
```{r}
str_split("apples, pears, and bananas", "")[[1]]
```

